Model Search in Relational Domains

### 2016-10-22 introduction.tex
edited intro

\section{Introduction}
%
%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
Static program analysis based on abstract 
interpretation~\cite{CC77} has been widely used to
verify certain classes of properties for safety-critical systems.  A 
standard static analyses~\cite{se2011} for program safety aims to compute 
invariants as fixed-points of abstract transformers.  Most lattice structure 
used in static analysis have meet and join operations.  The meet operation 
precisely model conjunction, and the join operation over-approximate 
disjunction. Thus, the precision loss in static analysers is often 
due to imprecise join operations. The loss in precision is overcome 
by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration 
behaviour. 
%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
On the other hand, the combination of precise conjunction in a partial
assignments domain with learning allows Conflict Driven Clause
Learning (CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to
reason about disjunction without enumerating cases.  Silva
et. al.~\cite{tacas12, sas12, dhk2013-popl} presented a formal recipe
for applying abstract interpretation to solve the satisfiability
problem.  \cite{dhk2013-popl} provide a generalisation of
CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning}
(ACDCL), in terms of deduction and abduction transformers,
fixed-point, interpolation and extrapolation operators on lattices.
The partial assignments in the CDCL algorithm are an abstract domain,
unit rule is the best abstract transformer, Boolean Constraint
Propagation (BCP) computes a greatest fixed-point over the abstract
domain, decision is dual widening and clause learning can be viewed as
synthesising an abstract transformer for negation.  
%Thus, ACDCL lifts learning techniques in SAT solver to operate on a non-distributive lattice -- thus equipping the analysis with negation. 
This enables the
analyser to refine the analysis and prevent enumeration behaviour.

ACDCL has so far been instantiated as a floating-point decision
procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser
that only uses interval constraints~\cite{tacas12}.  Hence, the
analysis is limited to bounds checking in simple numerical programs.
However, verification of certain class of programs require more
expressive domains like octagons or equalities or polyhedra.  These
domains capture interesting properties of programs which intervals can
not express.  In this paper, we identify specific properties of the
CDCL algorithm that are necessary to lift the propositional skeleton
of CDCL to relational abstract domains such as template polyhedra
domains~\cite{sriram} (octagons, zones and equalities).
To this end, we present a characterisation of abstract model search
procedure and conflict analysis procedure that lift the abstract
deduction transformer and the conflict analysis procedure to
relational lattice structures.

>>>>

\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
%
Static program analysis based on abstract interpretation~\cite{CC77} has
been widely used to verify of properties of safety-critical systems. 
Standard static analyses~\cite{se2011} aim to compute program invariants as
fixed-points of abstract transformers.  Abstract states are chosen from a
lattice that has meet and join operations.  The meet precisely models set
intersection (or conjunction, taking a logical view), and the join
over-approximates set union (or disjunction).  The over-approximation in the
join operation is a key cause of loss of precision, which yields false
alarms.  Standard means to address false alarms in static analysis are
richer abstract domains or equipping the analysis with more precise set
unions -- which is often very expensive since the analysis exhibits case
enumeration behaviour.

%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
%
On the other hand, propositional SAT solvers are known to be highly
effective for solving complex propositional formulae, and are frequently
able to avoid enumerating the exponential number of cases.  The impressive
performance of modern solvers is credited to well-tuned decision heuristics
and clause learning algorithms in modern solvers.  Collectively, these
algorithms are referred to as \emph{Conflict Driven Clause Learning}
(CDCL)~\cite{cdcl}. The idea to lift these algorithms from the Boolean
domain to richer, arithmetic domains suggests itself.

Abstract Conflict Driven Clause Learning} (ACDCL)~\cite{dhk2013-popl} is one
such generalisation of CDCL.  The partial satisfying assignments in CDCL
algorithm are generalized to be chosen from a given abstract domain.  The
unit rule is the best abstract transformer, Boolean Constraint Propagation
(BCP) computes a greatest fixed-point over the abstract domain, decision is
dual widening and clause learning can be viewed as synthesising an abstract
transformer for negation.  This enables an ACDCL-based analyser to refine an
imprecise analysis and thus prevents enumeration behaviour.  ACDCL has so
far been instantiated as a decision procedure for floating-point
arithmetic~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser that
reasons using interval~\cite{tacas12}.  Hence, the existing ACDCL instances
are limited to bounds checking in numerical programs.

Evidently, the verification of many programs requires more expressive
domains like octagons or equalities or polyhedra.  These domains capture
interesting properties of programs that intervals can not express.  In this
paper, we identify specific modifications to the CDCL algorithm that are
necessary to lift propositional CDCL to relational abstract domains such as
octagons, zones and equalities over numerical variables.  To this end, we
present a characterisation of an abstract model search procedure and a
conflict analysis procedure that lift the abstract deduction transformer and
the conflict analysis procedure to relational lattice structures.
